Load packages
```{r}
library(tidyverse)
library(base)
library(patchwork)
library(scales)
library(ggridges)
library(patchwork)
library(ggseqlogo)
library(table1)
```


Clean data

```{r}
#copy paste
rhc_clean <- rhc_data |> 
  bind_cols() |> 
  as_tibble() 

# convert death sting to binary: 1 = dead 0 = alive
rhc_clean |> mutate(death = case_when(
  death == "No" ~ 0,
  death == "Yes" ~ 1))

```
Load packages
```{r}
library(tidyverse)
library(base)
library(patchwork)
library(scales)
library(ggridges)
library(patchwork)
library(ggseqlogo)
library(table1)
```


Clean data

```{r}
#copy paste
rhc_clean <- rhc_data |> 
  bind_cols() |> 
  as_tibble() 

# convert death string to binary: 1 = dead 0 = alive
rhc_clean <- rhc_clean |> 
  mutate(death = case_when(
  death == "No" ~ 0,
  death == "Yes" ~ 1))

#round age to the nearest whole number
rhc_clean <- rhc_clean |> mutate("age" = round(age, 0)) 


## Delete all attributes when we know which ones we do not use

# Write the data to disk
write_csv(x = rhc_clean,
          file = "data/rhc_clean.csv.gz")













