## Description of the data
# Load packages
```{r}
library(tidyverse)
library(base)
library(patchwork)
library(scales)
library(ggridges)
library(patchwork)
library(ggseqlogo)
library(dplyr)
library(table1)
```

## copypaste your code snippets that fit for the description Dont delete your original document until we are completely finished, but rename it as __done._DONE Fks. mine is Josefine_DONE.qmd. This way we have a backup if something fucks up

##only write under your own name :)) 


Here we will describe the dataset:

- PCA on dataset of all the variables we chose to include to find what variables actually influence the decition of introducitn catherization
- divide into two catagories, soscioeconomic factors and physical factors: Find when a patient recieves RHC based on different diseases. Find how well the treatment works in terms based on death rate. Does income, edu, insurance, race, (sex?), influence when a patient actually recieves the treatment?

#Josefine

- cat1 and cat2 are primary and secondary disease category
```{r}
rhc_data |> group_by(cat1,cat2) |> summarise(n = n())
```

```{r}
rhc_data |> select(ends_with("hx")) |> summarise(n =n())
```
 
Table1 summary:
```{r}
T1 <- rhc_clean |> mutate(sex = factor(sex),
                    swang1 = factor(swang1),
                    death = factor(x = death, levels = c(0,1), c("Alive","Dead"))) |> 
  table1(x = formula(~ sex + age + race + swang1 | death),
         data = _) 
T1

```


  

#Emilie




#Sahand




#Judith




#Felix














