
```{r}
library(tidyverse)
library(base)
library(patchwork)
library(scales)
library(ggridges)
library(patchwork)
library(ggseqlogo)
library(dplyr)
library(table1)
```


The analysis explores the correlation between illnesses, Right Heart Catheterization (RHC) treatment, and the Acute Physiology Score (APS) concerning mortality. Various plots are generated to create an understanding of the dataset, aiming to reveal potential relationships among the mentioned attributes. These visual representations serve as a tool to discover whether specific factors are associated with death.

This plot shows all the diseases and how many people have it = 1 or dont have the disease = 0.
```{r}
subHX <- rhc_clean |> select(ends_with("hx"))

subHX <- subHX |> pivot_longer(everything(),
                      names_to = "variables",
                      values_to = "values")

subHX |> mutate(values = factor(values)) |> 
  ggplot(aes(x = variables, fill = values)) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  geom_bar(position = "dodge")

```

We begin by exploring if the number of comorbidity diseases influence the risk of death.
```{r}
rhc_aug |>
  ggplot(aes(x = comorbidities, fill=factor(death))) +
  geom_bar(position = "stack") +
  labs(title = "Risk of dying by number of comorbidities",
       x = "Number of comorbidities",
       y = "Number of patients") +
  scale_fill_manual(name= "Death", values = c("0" = "blue", "1" = "red")) +
  theme_minimal()
```


Now we want to include RHC treatment to see if this, together with comorbidities will influence the death rate. 
```{r}
rhc_aug |> mutate(swang1 = factor(x = swang1, levels = c(0,1), c("no RHC","RHC"))) |> 
  ggplot(aes(x = comorbidities, fill=factor(death))) +
  facet_wrap(~swang1)+
  geom_bar(position = "stack") +
  labs(title = "Risk of dying by number of comorbidities",
       x = "Number of comorbidities",
       y = "Number of patients") +
  scale_fill_manual(name= "Survival", values = c("0" = "#619CFF", "1" = "#F8766D"), labels= c('Dead', 'Alive')) +
  theme_minimal()
```


This plot shows all individuals who survided and died with and without the RHC treatment. 
```{r}

rhc_clean |> mutate(swang1 = factor(x = swang1, levels = c(0,1), c("no RHC","RHC")),
                    death = factor(x = death, levels = c(0,1), c("alive","dead"))) |> 
  count(death, swang1) |> 
  ggplot(aes(x = swang1, y = n, fill = death))+
  geom_col(alpha = 1,
           color="pink") +
  labs(x = 'Treatment type',
       y = 'n') +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

We investigate how many survive and die depending on their APS score. APS score predicts death by taking both laberatory values and patient signs into account. 
```{r}
rhc_aug |> 
  ggplot(aes(x = aps1, fill=factor(death))) +
  geom_bar(position = "stack") 
  
```
And then we make the x-axis an interval to bin the observations.
```{r}
rhc_aug |> 
  ggplot(aes(x = aps1_Interval, fill=factor(death))) +
  geom_bar(position = "stack") 

```




